Oral anticoagulants

### 2021-01-12 r_notebooks/exploring_oral_anticoagulants_war_prop.Rmd
Removed neur_net, typos

# Modifying functions to model log(n) as using proportion data
regionalCV <- function(data, cv_dist = 6, init = 36, step = 1){

  fableModels <- function(data){
    if(!is_tsibble(data)){data <- as_tsibble(data, index = datename)}

    data %>%
      mutate(datename = yearmonth(datename)) %>%
      model(trend_model1 = TSLM(log(n) ~ trend()),#time series linear model
            trend_model2 = TSLM(log(n) ~ trend() + season("year")),
            ets1 = ETS(log(n) ~ trend()),
            ets2 = ETS(log(n) ~ trend() + season("A")),
            ets3 = ETS(log(n) ~ trend() + season("M")),
            arima = ARIMA(log(n), stepwise = FALSE, approximation = FALSE),
            neur_net = NNETAR(log(n)),
            comb1 = combination_model(TSLM(log(n) ~ trend()),
                                      ETS(log(n) ~ trend())),
            comb2 = combination_model(TSLM(log(n) ~ trend() + season("year")),
                                      ETS(log(n) ~ trend() + season("A"))),
            stl_dcmp1 = decomposition_model(STL(log(n) ~ trend(),
                                                iterations = 1000,
                                                robust = TRUE),
                                            SNAIVE(season_adjust)),
            stl_dcmp2 = decomposition_model(STL(log(n) ~ trend() + season(),
                                                iterations = 1000,
                                                robust = TRUE),
                                            SNAIVE(season_adjust)),
            s_naive = SNAIVE(log(n)))
  }

@@

              tibble(holiday = "easter", # Sunday
                     ds = as.Date(c("2016-03-27", "2017-04-16", "2018-04-01",
                                    "2019-04-21", "2020-04-12", "2021-04-04")),
                     lower_window = -2,
                     upper_window = 1))

@@

    data_tr %>%
      fableModels() %>%
      forecast(h = cv_dist) %>%
      accuracy(data_test)

  }

@@

@@

# ## Read from database
# con <- DBI::dbConnect(RSQLite::SQLite(), 
#                       "../OneDrive/1_data/gpextract/db-gpextract.sqlite")
# db.gpextract <- tbl(con, "gpextract")
# oac_p <- db.gpextract %>%
#   filter(bnfchem %in% oral_anticoagulants) %>%
#   collect() %>%
#   mutate(datename = dmy(paste0("28-", datename)))  # Add day component to date and convert to date-time object
# DBI::dbDisconnect(con)
# 
# ## Write to local file
# oac_p %>%
#   write_rds("../data/oral_anticoagulants.rds")

@@

oac_p <- read_rds("../data/oral_anticoagulants.rds")

## Sort into monthly count data
oac_p_monthly_count <- 
  oac_p %>% 
  group_by(datename, bnfchem) %>% 
  summarise(n = sum(Quantity)) %>% 
  mutate(n = 100 * n / sum(n), # make it a percentage rather than proportion
         log_n = log(n)) %>% 
  filter(bnfchem == warf) %>% 
  ungroup() %>%
  as_tsibble(index = datename)

oac_p_monthly_quantity_hb <- 
  right_join(oac_p, select(gp, c(practiceid, hboard)), 
             by = c("PracticeID" = "practiceid")) %>%
  propRegional(region = hboard, datename = datename, unit = Quantity,
               bnf = warf)

oac_p_monthly_quantity_cluster <- 
  right_join(oac_p, select(gp, c(practiceid, cluster)), 
             by = c("PracticeID" = "practiceid")) %>%
  propRegional(cluster, datename, Quantity, bnf = warf)

oac_p_monthly_quantity_gp <- 
  propRegional(oac_p, PracticeID, datename, Quantity, bnf = warf)

@@

```{r oac_p-monthly-ts-plot, fig.cap="Total quantity of combined oral anticoagulants prescribed per month."}
oac_p %>% 
  group_by(datename) %>% 
  summarise(n = sum(Quantity)) %>% 
  tsibble() %>% 
  autoplot(n) +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

```{r oac-p-monthly-ts-plot, fig.cap="Total quantity of combined oral anticoagulants prescribed per month."}
oac_p_monthly_count %>% 
  select(datename, n) %>% 
  autoplot() +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

@@

* 724.044 seconds with multisession prophetModels()

```{r}
system.time(oac_p_hb_fcsts <- regionalJointFcsts(oac_p_monthly_quantity_hb_pre_cv19))

oac_p_hb_fcsts %>%
  write_rds("../data/oac_p_hb_fcsts.rds")

@@

* 194.071 seconds 2020-12-10

@@

* 5514.279 seconds

@@

* 1266.126 seconds 2020-12-10

>>>>

# Modifying functions to model log(n) as using proportion data
fableModels <- function(data){
  if(!is_tsibble(data)){data <- as_tsibble(data, index = datename)}

  data %>%
    mutate(datename = yearmonth(datename)) %>%
    model(trend_model1 = TSLM(log(n) ~ trend()),#time series linear model
          trend_model2 = TSLM(log(n) ~ trend() + season("year")),
          ets1 = ETS(log(n) ~ trend()),
          ets2 = ETS(log(n) ~ trend() + season("A")),
          ets3 = ETS(log(n) ~ trend() + season("M")),
          arima = ARIMA(log(n), stepwise = FALSE, approximation = FALSE),
          comb1 = combination_model(TSLM(log(n) ~ trend()),
                                    ETS(log(n) ~ trend())),
          comb2 = combination_model(TSLM(log(n) ~ trend() + season("year")),
                                    ETS(log(n) ~ trend() + season("A"))),
          stl_dcmp1 = decomposition_model(STL(log(n) ~ trend(),
                                              iterations = 1000,
                                              robust = TRUE),
                                          SNAIVE(season_adjust)),
          stl_dcmp2 = decomposition_model(STL(log(n) ~ trend() + season(),
                                              iterations = 1000,
                                              robust = TRUE),
                                          SNAIVE(season_adjust)),
          s_naive = SNAIVE(log(n)))
}

@@

            tibble(holiday = "easter", # Sunday
                   ds = as.Date(c("2016-03-27", "2017-04-16", "2018-04-01",
                                  "2019-04-21", "2020-04-12", "2021-04-04")),
                   lower_window = -2,
                   upper_window = 1))
regionalCV <- function(data, cv_dist = 6, init = 36, step = 1){

@@

    data_tr %>%
      fableModels() %>%
      forecast(h = cv_dist) %>%
      accuracy(data_test)
  }

@@

regionalJointFcsts <- function(data, h ="5 months"){
  regions <- unique(data$regional_unit)
  plan(multisession)
  fable_fit <-
    future_map_dfr(.x = regions,
                   .f = ~fableModels(filter(data, regional_unit ==.x)),
                   .options = furrr_options(seed = TRUE))
  message("fable fits complete")

  fable_fcst <- fable_fit %>%
    forecast(h = h) %>%
    as_tibble()
  message("fable forecasts complete")

  prophet_fit <-
    future_map_dfr(.x = regions,
                   .f = ~prophetModels(filter(data, regional_unit ==.x)),
                   .options = furrr_options(seed = TRUE))
  message("prophet fits complete")

  prophet_fcst <-
    forecast(prophet_fit, h = h) %>%
    as_tibble()
  # Some prophet forecasts are giving daily forecasts. We want monthly forecasts.
  # This may also be influencing accuracy

  message("prophet forecasts complete")
  plan(sequential)
  # prophetModels now being run in parallel. Needed to reinstall prophet from
  # source

  # joint_fit <-
  #   left_join(fable_fit, prophet_fit, by = "regional_unit")
  joint_fcst <-
    bind_rows(fable_fcst, prophet_fcst) %>%
    mutate(lb_95 = quantile(n, .025),
           lb_80 = quantile(n, .1),
           med = median(n),
           ub_80 = quantile(n, .9),
           ub_95 = quantile(n, .975),
           sd = distributional::variance(n)^2,
           dist = n,
           n = NULL)

  return(joint_fcst)
}

@@

# ## Read from database
# con <- DBI::dbConnect(RSQLite::SQLite(), 
#                       "../OneDrive/1_data/gpextract/db-gpextract.sqlite")
# db.gpextract <- tbl(con, "gpextract")
# oac <- db.gpextract %>%
#   filter(bnfchem %in% oral_anticoagulants) %>%
#   collect() %>%
#   mutate(datename = dmy(paste0("28-", datename)))  # Add day component to date and convert to date-time object
# DBI::dbDisconnect(con)
# 
# ## Write to local file
# oac %>%
#   write_rds("../data/oral_anticoagulants.rds")

@@

oac <- read_rds("../data/oral_anticoagulants.rds")

## Sort into monthly count data
oac_p_monthly_count <- 
  oac %>% 
  group_by(datename, bnfchem) %>% 
  summarise(n = sum(Quantity)) %>% 
  mutate(n = 100 * n / sum(n), # make it a percentage rather than proportion
         log_n = log(n)) %>% 
  filter(bnfchem == warf) %>% 
  ungroup() %>%
  as_tsibble(index = datename)

oac_p_monthly_quantity_hb <- 
  right_join(oac, select(gp, c(practiceid, hboard)), 
             by = c("PracticeID" = "practiceid")) %>%
  propRegional(region = hboard, datename = datename, unit = Quantity,
               bnf = warf)

oac_p_monthly_quantity_cluster <- 
  right_join(oac, select(gp, c(practiceid, cluster)), 
             by = c("PracticeID" = "practiceid")) %>%
  propRegional(cluster, datename, Quantity, bnf = warf)

oac_p_monthly_quantity_gp <- 
  propRegional(oac, PracticeID, datename, Quantity, bnf = warf)

@@

```{r oac_p-monthly-ts-plot, fig.cap="Total quantity of combined oral anticoagulants prescribed per month."}
oac %>% 
  group_by(datename) %>% 
  summarise(n = sum(Quantity)) %>% 
  tsibble() %>% 
  autoplot(n) +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

```{r oac-p-monthly-ts-plot, fig.cap="Warfarin prescribing as a percentage of all oral anticoagulants prescribed per month."}
oac_p_monthly_count %>% 
  select(datename, n) %>% 
  autoplot() +
  geom_vline(xintercept = dmy("23-March-2020"), colour = "red") # add line to mark the start of lockdown
```

@@

* 724.044 seconds with multisession prophetModels()
* 418.410 seconds with neur_net removed 2021-01-11

```{r}
plan(multisession)
system.time(oac_p_hb_fcsts <- regionalJointFcsts(oac_p_monthly_quantity_hb_pre_cv19))
plan(sequential)
oac_p_hb_fcsts %>%
  write_rds("../data/oac_p_hb_fcsts.rds")

@@

* 194.071 seconds 2020-12-10
* 71.559 seconds 2021-01-11

@@

* 5514.279 seconds
* 2059.159 seconds 2021-01-11

@@

* 1266.126 seconds 2020-12-10
* 590.595 seconds 2021-01-11
